---
output:
  html_document:
    css: C:/ALR/Models/custom.css
    highlight: textmate
    theme: null
---

# conteStreamflow Vignette: Data retrieval and Preparation
## Data prep for Massachusetts are monthly, seasonal, and annual data

### load packages
```{r load pkg, message=FALSE, warning=FALSE}


library(conteStreamflow)

```

### Set up cache and set local dir
```{r set up cache}

#set up cache
cache.setup( cache.dir="c:/alr/models/cache", quiet=T )
cache.set( cache.dir="c:/alr/models/cache" )

#local drive, for saving flow, weather, gage, and merged results
local.dir <- "c:/alr/models/conteStreamflow/vignettes/temp"
setwd(local.dir)

```

### Locate buffer file
```{r buffer file locations, cache=TRUE}
#this is the only input file needed, that is not accessed through a web service, or downloaded/cached from felek
buffer.file <- "C:/ALR/Models/conteStreamflow/data/mass_buffer_poly"
```

### Load gages using a buffer (polygon outline of area of interest)
```{r load gages, cache=TRUE}
g.spatial <- gage.retrieve( buffer.file=buffer.file, max.da.sqkm=50, min.da.sqkm=0 )

# number of records from initial retrieval
nrow(g.spatial)

```

### Place gages into NHDplus catchments to pair w/ stream reaches and basin characteristics
```{r place nhdplus, cache=TRUE}

g.spatial <- gage.place.nhdplus( g.spatial )


#all gages plotted to an nhdplus catchment
sum(is.na(g.spatial$FEATUREID))

```

### Load impoundments
```{r load impoundments, cache=TRUE}


#Load impoundment info
g.spatial<-impound.retrieve( gages.spatial=g.spatial, cols=char.columns.default( impound=T ) )


#Check if gages are missing impoundment data 
apply(g.spatial@data, MARGIN=2, FUN=function(x) sum(is.na(x)))
```

### Remove gages missing impoundment data, and filter out gages on impounded streams
```{r filter gages by impounded, cache=TRUE}

#Remove gages with dams or large on channel (open water) impoundments
#(and check #records before and after)
# Number of columns BEFORE gages w/ impoundments are removed
nrow(g.spatial)
g.spatial<-g.spatial[g.spatial$TNC_DamCount==0,]
g.spatial<-g.spatial[g.spatial$OnChannelWaterSqKM<.5,]
# Number of columns AFTER gages w/ impoundments are removed
nrow(g.spatial)


```


### Load basin characteristics
```{r load basin char, cache=TRUE}

#Let's use the default basin char columns
# Basin characteristic columns being used
print(char.columns.default()) 
#and pick a subset of those to log transform
log.cols <- c("Forest", "Herbacious","Agriculture","Developed","DevelopedNotOpen", "Impervious",  
               "CONUSOpenWater","CONUSWetland", "DrainageClass","HydrologicGroupAB","SurficialCoarseC","PercentSandy",
              "ReachElevationM", "BasinElevationM")

#Load basin characteristics
g.spatial<-char.retrieve( gages.spatial=g.spatial, cols=char.columns.default(), log.cols=log.cols )

```

### Remove gages/basins missing data
```{r gages missing data, cache=TRUE}


#See how many gages are missing each of the basin char
apply(g.spatial@data, MARGIN=2, FUN=function(x) sum(is.na(x)))


# One is missing ReachSlopePCNT.  Which gages?
g.spatial[is.na(g.spatial$ReachSlopePCNT),"station_nm"]


#Limit to gages that have all basin data
#(eliminate one missing ReachSlopePCNT, and recheck)
g.spatial<-g.spatial[!is.na(g.spatial$ReachSlopePCNT),]
apply(g.spatial@data, MARGIN=2, FUN=function(x) sum(is.na(x)))


# Total number of gages after removing those w/ missing basin char data
nrow(g.spatial)

#View sample of data
# Sample of gage data
head(g.spatial@data)
class(g.spatial)
str(g.spatial@data)  #structure of data slot, of sp object
```

### Look at mismatch of drainage area b/w NWIS gage info and NHDplus catchment
```{r da mismatch, cache=TRUE}

#since basin char are based on approximate mapping of gages to stream reaches using nhdplus catchments,
# some will be off/ will be mapped to the wrong stream sections or wrong stream/river
# look at mismatch between gage drainage area  and nhdplus drainage area

plot(g.spatial$da_sqkm,g.spatial$NHDplusTotDASqKM, main="Drainage areas of NWIS vs assigned NHDplus streams")
text(g.spatial$da_sqkm+1,g.spatial$NHDplusTotDASqKM-1,g.spatial$site_no,cex=.5)
abline( a=0, b=1, lty=2, col="red" )
abline( h=50, lty=2, col="black" )



# Will decide later what to do about this.  
# Not many problems in MA (after all, we have already removed dammed streams), but there are more problems elsewhere

```



### Place gages into weather grid
```{r place weather grid, cache=TRUE}
#Place gages into weather grid polygons 
g.spatial<-gage.place.weather( gages.spatial=g.spatial, plot=T )

#save rdata
setwd(local.dir)
save(g.spatial, file="gages_spatial.rdata")

#export shapefile
gage.export( g.spatial, shapefile.dir=local.dir )


```



### Load flow observations
```{r load flow, cache=TRUE}

#Load/calculate/aggregate flow data for seasonal and annual timesteps
q.matrices<-flow.retrieve( gages.spatial = g.spatial, 
                          periods = c("monthly","seasonal","annual") )

#View sample of flow data
# Sample flow data
str(q.matrices)
str(q.matrices[["monthly"]])
# Single gage, all columns
q.matrices[["monthly"]][360:370,2,]
# Single column, several gages
q.matrices[["monthly"]][360:370,1:10,1]


#save locally
setwd(local.dir)
save(q.matrices, file="q_matrices.rdata")
```


### Load weather observations
```{r load weather observations, cache=TRUE}


#Load/calculate/aggregate weather data for seasonal and annual timesteps
w.matrices <- weather.retrieve(gages.spatial=g.spatial, periods=c("monthly","seasonal","annual"))


# Structure and sample weather data
str(q.matrices)
str(q.matrices[["monthly"]])
# Single gage, all columns
w.matrices[["monthly"]][1:10,1,]
# Single column, several gages
w.matrices[["monthly"]][1:10,1:5,1]

#save locally
setwd(local.dir)
save(w.matrices, file="w_matrices.rdata")
```

### Merge flow, weather, and basin char data 
```{r merge data, cache=TRUE}

#merge
d <- data.merge( g.spatial, q.matrices, w.matrices, periods=c("monthly","seasonal","annual") )
#view sample and structure
str(d)
str(d[["monthly"]])
head(d[["monthly"]])


#save locally
setwd(local.dir)
d.annual <- d[["annual"]]
save(d.annual, file="d_annual.rdata")
d.seasonal <- d[["seasonal"]]
save(d.seasonal, file="d_seasonal.rdata")
d.monthly <- d[["monthly"]]
save(d.monthly, file="d_monthly.rdata")


```


### Footnotes: r version and packages used
```{r}
#objects in session
ls()

#session info and package versions
print(sessionInfo()) 

```

